---
title: "R Notebook"
output:
  html_document:
    code_folding: hide
---

## Author: OA Lab, NWFSC
## Title: Aquarium Water Chemistry Investigation: DO and Salinity corrections with new ramp period
## Date: July 2020


```{r 0.0 Libraries , echo=FALSE, results=FALSE}
#*********************************
##Libraries
#********************************* 
library(shiny)
library(tidyverse)
library(stringr)
library(readxl)
library(readr)
library(tidyr)
library(data.table)
library(lubridate)
library(violinmplot)
library(vioplot)
library(yarrr)
library(datapasta)
library(reprex)
library(miniUI)
library(gridExtra)
library(wql)
```



# Table of Contents
## Outline Current 2020.06.01

# 1. Setting the Working Directory
# 2. Calling & Reading in " dml "
# 3. Reformatting variables/vectors (Factors)
# 4. Creating dateTime objects
# 5. Creating Treatment Variables
# 6. Creating Night & Day Variables 
# 7. Salinity (4 New Vectors)
# 8. Creating Corrected Salinity Value 
# 9. DO
# 10. Filtering Water Chemistry Dataframe

# 1. Setting the Working Directory
1.0
Look for all files inside the Water Chemistry folder of the "06. MOATs replication verification" folder
```{r 1.0 Setting Working Directory, ECHO= FALSE, message=FALSE, warning=FALSE}
#*********************************
## 1.) Set working directory
#*********************************
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_knit$set(root.dir =
                       "/Users/katherinerovinski/GIT/NWFSC.MUK_KRL_SMR2019/06. MOATS replication verification/01a. WaterChemistry Data")
```

#### 1.1 Initial Working Directoy

Inside the working directory (folder) is 1 CSV files generated after combining 13CSV logs representing 24second observations. These CSV files were created using the moats graph app

#### 1.1.2 Link to Generating App

[Link to moats.Graph.app](http://github.com/pmcelhany/moatsGraphs.git)
Moving average window (nObs) set to 4. 

#### 1.1.3 Protocol with app

Every 17th observations was selected, the observations were 24second apart. 

#### 1.1.4 App data for 13 MOATS

1 CSVs per moats: M01, M02, M03, M04, M05, M06, M07, M08, M09, M10, M11, M12, M13. Files are also available on the OA Google drive. [OA Drive for MOATS data](https://drive.google.com/open?id=13a2hk1a9I9gRIgf2Xyl1dSYDf5vnyYvs)

### 1.2 Subsampling

For speed and ease of computations the dataframe was extermely subsampled. 
Instead of the standard "n17" used in the aquarium temperature investigation, this document subsampled at "n333".

#### 1.2.1 Subsampling explained by observations

All observations (per the 13 MOATS logs)
[1] 1137880   7 (6 seconds)
Number of observations after subsampling every 17th response/observation
[1] 66934     7 (24 seconds)
Number of observations after the super-subsampling, every 333rd observation
[1] 3417      7 (~30minutes) ***this won't detail the ramp


#### 1.2.2 Subsampling Determination

To examine the Day to Night ramp (D2Nramp) researcher used n17 subsampling group


# 2.0) Calling and Reading in dml

```{r 2.0 Calling and Reading in dml, ECHO= FALSE, message=FALSE, warning=FALSE, }
#*********************************
## 2. Calling & Reading in " dml " 
#*********************************

# Super Speedy subsampled data
dml <- read.csv(file = "M01thruM13moatslog_n17.csv", stringsAsFactors = FALSE)
dim(dml)

```

# 3.0) Reformatting variables/vectors (Factors)

Changing MOATs to Factors for the 13 different MOATs- these will be the discrete units for follow analysis

```{r Reformatting variables/vectors (Factors), ECHO= FALSE, message=FALSE, warning=FALSE}
#*********************************
## 3. Reformatting variables/vectors (Factors) 
#*********************************

## 3.3 Changing variables | 
dml$moats <- factor(dml$moats)
# Checking the names of the different levels
levels(dml$moats)
##checking the dataset, dimensions
dim(dml)
```


# 4.0) Creating dateTime objects  

```{r 4.0 establish the date time object of the CSV, ECHO= FALSE, message=FALSE, warning=FALSE}
dml$dateTime <- as.POSIXct(dml$dateTime, format="%Y-%m-%d %H:%M:%OS")
ReferenceTime <- as.POSIXct("2019-09-20 23:59:00")
class(ReferenceTime)
```


























.**************E*N*D*************# 
.*********************************
.# END OF SCRIPT | END OF DOCUMENT 
.*********************************


. ___________________8888,
. ____________________Y8888b,
. ___________________,oA8888888b,
. _____________,aaad8888888888888888bo,
. __________,d888888888888888888888888888b,
. ________,888888888888888888888888888888888b,
. _______d8888888888888888888888888888888888888,
. ______d888888888888888888888888888888888888888b
. _____d888888P'                    `Y888888888888,
. _____88888P'                    Ybaaaa8888888888l
. ___a8888'                      `Y8888P' `V888888
. __d8888888a                                `Y8888
. AY/'' `\Y8b                                 ``Y8b
. Y'      `YP                                    ~~