---
title: "R Notebook"
output: html_notebook
---

## Author: OA Lab, NWFSC
## Title: Aquarium Water Chemistry Investigation: DO and Salinity corrections with new ramp period
## Date: July 2020


```{r}
#*********************************
##Libraries
#********************************* 
library(shiny)
library(tidyverse)
library(stringr)
library(readxl)
library(readr)
library(tidyr)
library(data.table)
library(lubridate)
library(violinmplot)
library(vioplot)
library(yarrr)
library(datapasta)
library(reprex)
library(miniUI)
library(gridExtra)
library(wql)
```



# Table of Contents
## Outline Current 2020.06.01

# 1. Setting the Working Directory
# 2. Calling & Reading in " dml "
# 3. Reformatting variables/vectors (Factors)
# 4. Creating dateTime objects
# 5. Creating Treatment Variables
# 6. Creating Night & Day Variables 
# 7. Salinity (4 New Vectors)
# 8. Creating Corrected Salinity Value 
# 9. DO
# 10. Filtering Water Chemistry Dataframe


```{r}
#*********************************
## 1.) Set working directory
#*********************************
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_knit$set(root.dir =
                       "/Users/katherinerovinski/GIT/NWFSC.MUK_KRL_SMR2019/06. MOATS replication verification/01a. WaterChemistry Data")
```


