---
title: "Aquarium Water Chemistry Investigation: DO and Salinity corrections, Ramp Period highlights"
output:
  html_document:
    code_folding: hide
---

## Author: OA Lab, NWFSC
## Title: Aquarium Water Chemistry Investigation: DO and Salinity corrections with new ramp period
## Date: July 2020


```{r 0.0 Libraries , echo=FALSE, results=FALSE}
#*********************************
##Libraries
#********************************* 
library(shiny)
library(tidyverse)
library(stringr)
library(readxl)
library(readr)
library(tidyr)
library(data.table)
library(lubridate)
library(violinmplot)
library(vioplot)
library(yarrr)
library(datapasta)
library(reprex)
library(miniUI)
library(gridExtra)
library(wql)
```



# Table of Contents
## Outline Current 2020.06.01

# 1. Setting the Working Directory
# 2. Calling & Reading in " dml "
# 3. Reformatting variables/vectors (Factors)
# 4. Creating dateTime objects
# 5. Creating Treatment Variables
# 6. Creating Night & Day Variables 
# 7. Salinity (4 New Vectors)
# 8. Creating Corrected Salinity Value 
# 9. Creating Percent Dissoved Oxy Value 
# 10. Filtering Water Chemistry Dataframe
# 11. Averages
# 12. DO summary tables & plots
# 13. pH summary tables & plots




# 1. Setting the Working Directory
1.0
Look for all files inside the Water Chemistry folder of the "06. MOATs replication verification" folder


```{r, setup, include=FALSE }
#*********************************
## 1.) Set working directory
#*********************************
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_knit$set(root.dir =
                       "/Users/katherinerovinski/GIT/NWFSC.MUK_KRL_SMR2019/06_MOATS_replication_verification/01a_WaterChemistry_Data")

getwd()
setwd("/Users/katherinerovinski/GIT/NWFSC.MUK_KRL_SMR2019/06_MOATS_replication_verification/01a_WaterChemistry_Data")
getwd()
```





#### 1.1 Initial Working Directoy

Inside the working directory (folder) is 1 CSV files generated after combining 13CSV logs representing 24second observations. These CSV files were created using the moats graph app

#### 1.1.2 Link to Generating App

[Link to moats.Graph.app](http://github.com/pmcelhany/moatsGraphs.git)
Moving average window (nObs) set to 4. 

#### 1.1.3 Protocol with app

Every 17th observations was selected, the observations were 24second apart. 

#### 1.1.4 App data for 13 MOATS

1 CSVs per moats: M01, M02, M03, M04, M05, M06, M07, M08, M09, M10, M11, M12, M13. Files are also available on the OA Google drive. [OA Drive for MOATS data](https://drive.google.com/open?id=13a2hk1a9I9gRIgf2Xyl1dSYDf5vnyYvs)

### 1.2 Subsampling

For speed and ease of computations the dataframe was extermely subsampled. 
Instead of the standard "n17" used in the aquarium temperature investigation, this document subsampled at "n333".

#### 1.2.1 Subsampling explained by observations

All observations (per the 13 MOATS logs)
[1] 1137880   7 (6 seconds)
Number of observations after subsampling every 17th response/observation
[1] 66934     7 (24 seconds)
Number of observations after the super-subsampling, every 333rd observation
[1] 3417      7 (~30minutes) ***this won't detail the ramp


#### 1.2.2 Subsampling Determination

To examine the Day to Night ramp (D2Nramp) researcher used n17 subsampling group


# 2.0) Calling and Reading in dml
```{r 2.0 Calling and Reading in dml, ECHO= FALSE, message=FALSE, warning=FALSE, }
#*********************************
## 2. Calling & Reading in " dml " 
#*********************************

# Super Speedy subsampled data
dml <- read.csv(file = "M01thruM13moatslog_n17.csv", stringsAsFactors = FALSE)
dim(dml)


```



# 3.0) Reformatting variables/vectors (Factors)

Changing MOATs to Factors for the 13 different MOATs- these will be the discrete units for follow analysis

```{r Reformatting variables/vectors (Factors), ECHO= FALSE, message=FALSE, warning=FALSE}
#*********************************
## 3. Reformatting variables/vectors (Factors) 
#*********************************

## 3.3 Changing variables | 
dml$moats <- factor(dml$moats)
# Checking the names of the different levels
levels(dml$moats)
##checking the dataset, dimensions
dim(dml)
```


# 4.0) Creating dateTime objects  
Changes to the format of dates and times for observations

```{r 4.0 establish the date time object of the CSV, ECHO= FALSE, message=FALSE, warning=FALSE}
dml$dateTime <- as.POSIXct(dml$dateTime, format="%Y-%m-%d %H:%M:%OS")
ReferenceTime <- as.POSIXct("2019-09-20 23:59:00")
class(ReferenceTime)

# QA check
dim(dml)

```

# 5.0) Creating Treatment Variables  

#### 5.1 Three Treatments
Three treatments of this studies are identified as 
* "current" for the Current Conditions Treatment
* "hightemperature" for the High Temperature Conditions Treatment
* "allchange" for the All Change Conditions Treatment which incorporated both the high temperature conditions of the "hightemperature" treatment along with exposure to lower aquariun pH.   

```{r 5.1 Identifying treatments by moats, ECHO= FALSE, message=FALSE, warning=FALSE}
## 5.1 Identifying treatments by moats 
## establishing treatments
dml$treatment <- ""
dml$treatment[dml$moats == "M07" | dml$moats== "M10" | dml$moats== "M12"] <- "current"
dml$treatment[dml$moats == "M01"| dml$moats== "M06"] <- "hightemperature"
dml$treatment[dml$moats == "M02"| dml$moats== "M08" | dml$moats== "M13"] <- "allchange"
dml$treatment[dml$moats == "M03"| dml$moats == "M04" | dml$moats == "M09" | dml$moats == "M05" | dml$moats== "M11"] <- "broken_and_ambientbroken"
#verify that this new column has been created
names(dml)
#results should include:
#[1] "moats"        "dateTime"     "aTemperature" "sTemperature" "pH"          
#[6] "DO"           "salinity"     "treatment"  

# QA check
dim(dml)

```

#### 5.1.1 Ambient Treatment
Those MOATS aquarium systems that did not reach desired/programed conditions were deemed the "ambient" treatment. No temperature, DO, or pH conditions were programmed or alarmed. Flow control was governed by a solenoid valve to acheive flow/no-flow, day/night periods. 

MOATS inside the "Ambient" treatment included MOATs 04 and MOATs 05 both were shown to have faulty thermisters. 

Recorded conditions of what the animals experienced can't be guaranteed/confirmed. 

Without a record of aquarium settings MOATs 04 and MOAT 05 data was not included. 



# 6.) Creating Night and Day Periods 

#### 6.1 Narrative (Overall)
Creating a day and night variables 
Day and night periods will only refer to time under treatment as a way to exclude the acclimation period.
Day and night changed at about ~ 1230 on 05OCT19 
Treatment start date considered to begin Monday 23SEP19 at 1200pm
Krill Night Starts 1200 (~1230*) and ends 2100
Krill Days Starts 2101 and ends 1159 (~1229*) 
Interval 1 start 1200 23SEP19, end 1229 05OCT19
Interval 2 start 1230 05OCT19, end 2100 30OCT19

#### 6.2 New Column, New Variable in dml 

```{r 6.2 New Column, New Variable in dml}
## 6.2 New Column, New Variable in dml
#creating a new column, new variable "period"
dml$period <- ""
```

#### 6.3 Disassembling dateTime to create 2 new variables

```{r 6.3 Disassembling dateTime to create 2 new variables}
## 6.3 Disassembling dateTime to create 2 new variables
# Create new split date and time columns
dml$ObservationDate <- as.Date(dml$dateTime)
dml$ObservationTime <- format(as.POSIXct(dml$dateTime) ,format = "%H:%M:%S")
```

#### 6.4 Narrative about Intervals 
Interval 1
Interval Date Start  "2019-09-23"
Interval Date End    "2019-10-05"
Day Start Time       "21:01:00"
Day End Time         "11:29:00"
Night Start Time     "12:00:00"
Night End Time       "21:00:00"
DaytoNight Ramp(D2N) "11:30:00"
start
DaytoNight Ramp(D2N) "11:59:00"
stop
Other Time


Interval 2
Interval Date Start  "2019-10-05"
Interval Date End    "2019-10-30"
Day Start Time       "21:01:00"
Day End Time         "12:29:00"
Night Start Time     "12:30:00"
Night End Time       "21:00:00"
Other Time

#### 6.5 Period Assignments

```{r 6.5 Day / Night Assignments,  ECHO= FALSE, message=FALSE, warning=FALSE}
## 6.5 Day / Night Assignments 
# Using the "case_when" function in the tidyverse in the place of a loop
 
dml <- dml %>% mutate(period=case_when(
  (ObservationDate >= "2019-09-23") 
  & (ObservationDate <="2019-10-05") 
  & (ObservationTime >= "12:00:00") 
  & (ObservationTime <="21:00:00") ~"night",
  
  (ObservationDate >= "2019-10-05")
  & (ObservationDate <= "2019-10-30")
  & (ObservationTime >= "12:31:00") 
  & (ObservationTime <="21:00:00") ~"night",
  
  (ObservationDate >= "2019-09-23") 
  & (ObservationDate <="2019-10-05")
  & ((ObservationTime >= "21:01:00") 
     | (ObservationTime <="11:29:00")) ~"day",
  
  (ObservationDate >= "2019-10-05")
  & (ObservationDate <= "2019-10-30")
  & ((ObservationTime >= "21:01:00")
     | (ObservationTime <= "12:01:00")) ~"day",
  
  (ObservationDate >= "2019-09-23") 
  & (ObservationDate <="2019-10-05")
  & ((ObservationTime >= "11:30:00") 
     | (ObservationTime <="12:00:00")) ~"D2Nramp",
  
  (ObservationDate >= "2019-10-05")
  & (ObservationDate <= "2019-10-30")
  & ((ObservationTime >= "12:01:00")
     | (ObservationTime <= "12:30:00")) ~"D2Nramp",

  
  TRUE ~"other"
)

) 
```

#### 6.5.1 Quick Check on Period Generation

```{r 6.5.1 aTemperature Time Series Plots to Invesitgate Different Intervals}
#Quick check to see if periods were created

period.intervals <- ggplot(dml, aes(x=dateTime, y=aTemperature)) +
  geom_point(aes(colour=period, point=))   +
  ggtitle("aTemperature Time Series Plots to Invesitgate Different Intervals") +
  ylim (5.0, 15.00)

period.intervals
```

#### 6.5.1 Quick Check on Day to Night Ramp Verification
The ramp verification will be important to dissolved oxygen verification
Dissolved oxygen not yet shown. 
Recorded dissolved oxygen needs to be corrected with observed salinity. 

The following 2 plots are only present to visualize the different periods across treatments

```{r  6.5.1a Quick Check on Day to Night Ramp Verification}
ggplot(subset(dml, 
               period %in% ("D2Nramp")), 
        aes(x=dateTime, y=aTemperature)) + 
   geom_point(aes(colour=treatment, point=)) +
   ylim (5.0, 15.00) + 
   ggtitle("aTemperature Values, All MOATs, During Day to Night Ramp Period")

```



```{r 6.5.1b Quick Check on Day to Night Ramp Verification}

ggplot(subset(dml[dml$treatment == "allchange", ], 
             period %in% ("D2Nramp")), 
       aes(x=dateTime, y=aTemperature)) + 
  geom_point(aes(colour=moats, point=)) +
  ggtitle("Aquarium Temperatures, All Change, During Ramp Period")

```


#### 6.6 Removing the Acclimation Period 
The acclimation and the time after animals were removed from the system was designated other

```{r}

# Removing "other" period from day and night 
# not including acclimation period in this investigation

H2Ochemdf <- dml %>% filter(period != "other")
```

#### Water Chemistry Dataframe Created "H2Ochemdf"



# 7.) Bringing in Salinity   


```{r 7.0 Bringing in Salinity }
)

knitr::opts_chunk$set(echo = TRUE)

knitr::opts_knit$set(root.dir =
                       "/Users/katherinerovinski/GIT/NWFSC.MUK_KRL_SMR2019/06_MOATS_replication_verification/01a_WaterChemistry_Data")

getwd()
setwd("/Users/katherinerovinski/GIT/NWFSC.MUK_KRL_SMR2019/06_MOATS_replication_verification/01a_WaterChemistry_Data")
getwd()




all_salinity <- read.csv(file = "KRL19_salinityreadings_all.csv", stringsAsFactors = FALSE)
all_salinity$PSU_ObsDate <- ""#

all_salinity$Date <- as.POSIXct(all_salinity$Date, format= "%m/%d/%y")
all_salinity$PSU_ObsDate <- as.Date(all_salinity$Date)
```


# 8.0) Creating Corrected Salinity Value   

#### 8.1) Narrative: Creating Corrected Salinity Value from Point Measurments 
Notes from the 2020.05.29 office hours 
creating 4 new variables in Water Chemistry Dataframe
create a vector 181845  for the large n17 file
create a vector 9283 observations long for the sub-sub sample n333 file
logical vectors- boolean answers to those three below conditions
1 - the measurement per moats per day - MOATs and Data are both available
2 - the measurement (averaged across moats) - group by / summarize tools
3- (data gapped situation- no value matches) take the previous daily average based on observation  - lag? 
previous line on a dataframe ... dplyr tool ... 
 https://dplyr.tidyverse.org/reference/lead-lag.html
4 - "the winning value" case_when #corrected conductivity value


```{r 8.1) Creating Corrected Salinity Value}
# Salinity Values 
H2Ochemdf$PSUperMOATs <- ""
H2Ochemdf$PSUavgDaily <- ""
H2Ochemdf$PSUprevObs <- ""
H2Ochemdf$assumed_PSU <- ""
H2Ochemdf$Final_PSU <- ""
```


#### 8.1 Plots (timeseries verification)
#### Simple Plot 8.1.1
```{r 8.1.1 Plots (timeseries verification)}
# 8.1 Plots (timeseries verification)

p <- ggplot(H2Ochemdf, aes(x=ObservationDate, y=salinity))+
             geom_line() + 
             ylim (30.05, 27.00) 
p

```

#### Simple Plot 8.1.2 Salinity Values all Treatments, All MOATs, During Night Period

```{r 8.1.2 Salinity Values all Treatments, All MOATs, During Night Period}
ggplot(subset(H2Ochemdf, 
               period %in% ("night")), 
        aes(x=ObservationDate, y=salinity)) + 
        geom_point(aes(colour=period, point=)) +
        ggtitle("Salinity Constants all Treatments, All MOATs, During Night Period")
```

#### Simple Plot 8.1.3

```{r #### 8.1.3 Salinity Constants all Treatments, All MOATs, During Night Period }

ggplot(subset(H2Ochemdf, 
               period %in% ("night")), 
        aes(x=ObservationDate, y=salinity)) + 
        geom_point(aes(colour=moats, point=)) +
        ggtitle("Salinity Constants all Treatments, All MOATs, During Night Period")
```

#### 8.1.4 Salinity Constants, MOATs 01, During Night Period

```{r 8.1.4 Salinity Constants all Treatments, All MOATs, During Night Period}
ggplot(subset(H2Ochemdf[H2Ochemdf$moats == "M01", ], 
               period %in% ("day")), 
        aes(x=dateTime, y=salinity)) + 
   geom_point(aes(colour=moats, point=)) +
   ggtitle("Salinity Constant, MOATs 01, During Night Period")
```




#### 8.2 Practical Salinity Units - Measurements Per Moats per Day
These observations consist of salinity measurements taken per MOATS per day.
The observation were recorded per MOATS per day as the researcher took point measurements from the aquarium box while doing DO measurements. 

```{r 8.2 Practical Salinity Units - Measurements Per Moats per Day}
# Manual Readings
# Per MOAT measurement
H2Ochemdf$PSUperMOATs <- all_salinity$salinity[match
                                               (paste
                                                 (H2Ochemdf$ObservationDate,
                                                   H2Ochemdf$moats), 
                                                 paste
                                                 (all_salinity$PSU_ObsDate, 
                                                   all_salinity$moats))]
```

#### 8.3 Practical Salinity Units - Daily Avg Measurement 
Manual records sometimes have holes
Monday through Friday averages were recorded with greater fidelity than weekend averages. 


```{r 8.3 Practical Salinity Units - Daily Avg Measurement}
# All MOATs daily
H2Ochemdf$PSUavgDaily <- all_salinity$salinity[match
                                               (paste
                                                 (H2Ochemdf$ObservationDate,'All'), 
                                                 paste
                                                 (all_salinity$PSU_ObsDate, 
                                                   all_salinity$moats), 
                                                 nomatch = NA_character_)]
```


#### 8.4 Practical Salinity Units - Previous Daily Avg Measurement 
DO point measurements occured weekly. 
Salinity averages were typically recorded every day, Monday- Friday.
Some DO measurements were a week apart. 
The previous daily average was thought to be a better reflection than the last per Day, per MOATs salinity measurement.

#### 8.4.1 Last Observation Carried Forward Function
(Description)
Generic function for replacing each NA with the most recent non-NA prior to it.

This function was used to cover those gaps that occured most weekends.

```{r 8.4 Practical Salinity Units - Previous Daily Avg Measurement}
# Previous All MOATs daily
H2Ochemdf$PSUprevObs <- na.locf(H2Ochemdf$PSUavgDaily, na.rm = FALSE)
```


#### 8.5 Practical Salinity Units - Assumed Salinity Value 
Before the start of the study, a salinity of 28.8PSU was observed the week prior to the study's start

```{r 8.5 Practical Salinity Units - Assumed Salinity Value}
 
H2Ochemdf$assumed_PSU <- 28.8

```

#### 8.6 Practical Salinity Units - Final Salinity Value 
cycling between different salinity measurements required moving between character based and numeric based functions.

```{r 8.6a Review of Vectors, Creating Final_PSU value}

# 8.6a Review of Vectors
# Review of the three different salinity values
# H2Ochemdf$PSUperMOATs       measurement
#           PSUperMOATs       per MOATS per Day

## PSUperMOATs

# H2Ochemdf$PSUprevObs        Daily Average
#           PSUavgDaily       measurment across MOATs

## PSUavgDaily 

# H2Ochemdf$PSUprevObs        Previous Daily 
#           PSUprevObs        Average to cover those gaps

## PSUprevObs 

# Sailinity Assumption        28.8
#           Dates w/o         Dates without measurment 23~30SEP20
28.8


# Final Salinity
H2Ochemdf$Final_PSU <- 0.0

# H2Ochemdf$PSUperMOATs <- ""
# H2Ochemdf$PSUavgDaily <- ""
# H2Ochemdf$PSUprevObs <- ""
# H2Ochemdf$assumed_PSU <- ""

H2Ochemdf$Final_PSU <- as.numeric(case_when(
  H2Ochemdf$PSUperMOATs != 'NA' ~ H2Ochemdf$PSUperMOATs,
  H2Ochemdf$PSUavgDaily != 'NA' ~ H2Ochemdf$PSUavgDaily,
  H2Ochemdf$PSUprevObs != 'NA' ~ H2Ochemdf$PSUprevObs,
  TRUE ~ as.character(H2Ochemdf$assumed_PSU),
))

```


#### 8.7 Plots of Salinity 

#### 8.7a Plots of Salinity, Salinity Observations per day, MOATs 01, During Day Period
```{r 8.7 Plots of Salinity Salinity Observations per day, MOATs 01, During Day Period}
ggplot(subset(H2Ochemdf[H2Ochemdf$moats == "M01", ], 
               period %in% ("day")), 
        aes(x=dateTime, y=PSUperMOATs)) + 
   geom_point(aes(colour=moats, point=)) +
   ggtitle("Salinity Observations per day, MOATs 01, During Day Period")
```

####8.7b Salinity Observations per day, All Change Treatment, Day Period"

```{r}

ggplot(subset(H2Ochemdf[H2Ochemdf$treatment == "allchange", ], 
               period %in% ("day")), 
        aes(x=dateTime, y=PSUperMOATs)) + 
   geom_point(aes(colour=moats, point=)) +
   ggtitle("Salinity Observations per day, All Change Treatment, Day Period")

```


####8.7b.1 Salinity Observations per day, Final PSU Value, All Change Treatment, Day Period"

```{r}

ggplot(subset(H2Ochemdf[H2Ochemdf$treatment == "allchange", ], 
               period %in% ("day")), 
        aes(x=dateTime, y=Final_PSU)) + 
   geom_point(aes(colour=moats, point=)) +
   ggtitle("Salinity Observations per day, Final PSU Value, All Change Treatment, Day Period")

```






#### 8.7c Salinity Observations per day, Current Conditions Treatment, Day Period
```{r}
ggplot(subset(H2Ochemdf[H2Ochemdf$treatment == "current", ], 
               period %in% ("day")), 
        aes(x=dateTime, y=PSUperMOATs)) + 
   geom_point(aes(colour=moats, point=)) +
   ggtitle("Salinity Observations per day, Current Conditions Treatment, Day Period")
```


####8.7c.1 Salinity Observations per day, Final PSU Value, Current Conditions Treatment, Day Period"

```{r}

ggplot(subset(H2Ochemdf[H2Ochemdf$treatment == "current", ], 
               period %in% ("day")), 
        aes(x=dateTime, y=Final_PSU)) + 
   geom_point(aes(colour=moats, point=)) +
   ggtitle("Salinity Observations per day, Final PSU Value, Current Conditions Treatment, Day Period")

```




#### 8.7d Salinity Observations per day, High Temperature Conditions Treatment, Day Period
```{r}
ggplot(subset(H2Ochemdf[H2Ochemdf$treatment == "hightemperature", ], 
               period %in% ("day")), 
        aes(x=dateTime, y=PSUperMOATs)) + 
   geom_point(aes(colour=moats, point=)) +
   ggtitle("Salinity Observations per day, High Temperature Conditions Treatment, Day Period")
```

####8.7d.1 Salinity Observations per day, Final PSU Value, High Temperature Conditions Treatment, Day Period"

```{r}

ggplot(subset(H2Ochemdf[H2Ochemdf$treatment == "hightemperature", ], 
               period %in% ("day")), 
        aes(x=dateTime, y=Final_PSU)) + 
   geom_point(aes(colour=moats, point=)) +
   ggtitle("Salinity Observations per day, Final PSU Value, High Temperature Conditions Treatment, Day Period")

```




```{r}


ggplot(H2Ochemdf, aes(treatment, Final_PSU)) +
            geom_jitter(color = "grey") +
            geom_boxplot(notch = TRUE, outlier.shape = NA, colour = "black") +
            facet_wrap(~period) +
            ggtitle("All Treatments, Final Salinity Values (PSU) Across Periods") +
            theme_bw() 


```



# 9.0) Creating Percent Dissoved Oxy Value 

The Library wql is required for the following chunk 
The package name stands for ``water quality'' and reflects the # original focus on time series data for physical and chemical # properties of water, as well as the biota. Intended for # programs that sample approximately monthly, quarterly or # annually at discrete stations, a feature of many legacy data # sets. Most of the functions should be useful for analysis of # similar-frequency time series
regardless of the subject # matter.

#### 9.1) Creating Percent Dissolved Oxy Value - creating variables  

```{r 9.1 Creating Percent Dissolved Oxygen Value,  ECHO= FALSE, message=FALSE, warning=FALSE}

#assumption with the 28.8 standard salinity reading
H2Ochemdf$percentDOassumpt <- ""
H2Ochemdf$percentDOassumpt <- as.numeric(H2Ochemdf$percentDOassumpt)

H2Ochemdf$assumedSatDOmg <- ""
H2Ochemdf$assumedSatDOmg <- as.numeric(H2Ochemdf$assumedSatDOmg)

# the percent DO
H2Ochemdf$percentDO <- "" 
H2Ochemdf$percentDO <- as.numeric(H2Ochemdf$percentDO)


# Observed / Measured Salinity readings informed answers
H2Ochemdf$obseveredSatDOmg <- "" 
H2Ochemdf$obseveredSatDOmg <- as.numeric(H2Ochemdf$obseveredSatDOmg)


H2Ochemdf$actualDOmg <- ""
H2Ochemdf$actualDOmg <- as.numeric(H2Ochemdf$actualDOmg)

# Review of Values
H2Ochemdf$percentDOassumpt
H2Ochemdf$assumedSatDOmg
H2Ochemdf$percentDO
H2Ochemdf$obseveredSatDOmg
H2Ochemdf$actualDOmg


```




#### 9.2) Creating Percent Dissoved Oxy Value - assumed DO saturation variable creation


```{r 9.2) Creating Percent Dissoved Oxy Value - assumed DO saturation variable creation}

# saturated mg/L DO at obseved temperature and assumed salinity
# the oxySol() function is form the wql package
H2Ochemdf$assumedSatDOmg <- oxySol(H2Ochemdf$sTemperature, 
                                   H2Ochemdf$assumed_PSU)


SatDOplot <- ggplot(H2Ochemdf, aes(x=ObservationDate, 
                                   y=assumedSatDOmg))+
  geom_point(aes(colour=treatment, point=)) +
  ggtitle("Assumed Saturation of Dissolved Oxygen Across all MOATS")

SatDOplot


```


#### 9.3) Creating Percent Dissoved Oxy Value - Percent DO (with assumptions)


```{r 9.3 Creating Percent Dissoved Oxy Value - Percent DO (with assumptions) }
# Back calculated fraction DO as reported by the oxygen sensor

H2Ochemdf$percentDOassumpt <- ""
H2Ochemdf$percentDOassumpt <- as.numeric(H2Ochemdf$percentDOassumpt)
H2Ochemdf$percentDOassumpt <- H2Ochemdf$DO / H2Ochemdf$assumedSatDOmg


p5 <- ggplot(H2Ochemdf, aes(x=dateTime, y=percentDOassumpt)) +
  geom_point(aes(colour=period, point=))   +
  ggtitle("Assumed Percent DO (All Treatments & All MOATs) Colored by Period")

p5
```

#### 9.3a Filtering to Display ramp DO activity

```{r 9.3a Filtering to display ramp DO activity}

f.H2Ochemdf.ramp = filter(H2Ochemdf,
  !moats %in% c('M03', "M04", "M05", "M11") & 
  treatment %in% c("current", "allchange", "hightemperature") &
  period != "night")

f.H2Ochemdf.ramp$moats <- droplevels(f.H2Ochemdf.ramp$moats)
f.H2Ochemdf.ramp$treatment <- factor(f.H2Ochemdf.ramp$treatment)


p5 <- ggplot(f.H2Ochemdf.ramp, aes(x=dateTime, y=percentDOassumpt)) +
  geom_point(aes(colour=period, point=))   +
  facet_wrap(~treatment)
  ggtitle("Assumed Percent DO (All Treatments & All MOATs) Ramping Up From Day to Night")

p5



```

```{r}

#*********************************
## 9.5) Creating Percent Dissoved Oxy Value - Observed Saturated DOmg (measured salinity)
#*********************************

# #satured mg/L at observed temperature and observed (not assumed) salinity
H2Ochemdf$sTemperature <- as.numeric(H2Ochemdf$sTemperature)

H2Ochemdf$obseveredSatDOmg <- oxySol(H2Ochemdf$sTemperature, H2Ochemdf$Final_PSU)

```





```{r}
#*********************************
## 9.6) Creating Percent Dissoved Oxy Value - Observed/Actual DO (in mg) Saturated DOmg (measured salinity)
#*********************************
# actual DO mg at observed temperature and salinity
H2Ochemdf$percentDO <- H2Ochemdf$DO / H2Ochemdf$assumedSatDOmg

H2Ochemdf$actualDOmg <- H2Ochemdf$percentDO * H2Ochemdf$obseveredSatDOmg


actualDOmgPLOT <- ggplot(H2Ochemdf, aes(x=dateTime, y=H2Ochemdf$actualDOmg)) +
  geom_point(aes(colour=period, point=))   +
  ggtitle("Actual DO (mg/L) (All Treatments & All MOATs) Colored by Period")

actualDOmgPLOT
```





# 10.0) Filtering Data
```{r 10.0 Filtering Data across all }
# # 10.0) Filtering Data

H2Ochemdf <- H2Ochemdf  %>% filter(!moats %in% c("M03", "M04", "M05", "M11")) %>%
  filter(sTemperature>= 5 & sTemperature<=30) %>%
  filter(treatment %in% c("current", "allchange", "hightemperature")) %>%
  filter(period != "other")


write.csv(H2Ochemdf, "2020.07.24_H2Ochemdf.csv")

```


```{r 10.1) Filtering Data- dropping lvls to clean graphs}
## 10.1) Filtering Data- dropping lvls to clean graphs
#*********************************


filteredFrame = filter(H2Ochemdf,
                       !moats %in% c('M03', "M04", "M05", "M11") & 
                         (aTemperature>= 5 & aTemperature<=30) &
                         treatment %in% c("current", "allchange", "hightemperature") &
                         period != "other")

#|- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - |

## 10.1a Dropping levels and factors
filteredFrame$moats <- droplevels(filteredFrame$moats)
filteredFrame$treatment <- factor(filteredFrame$treatment)


```


# 11.0) DO averages- days and nights

#### 11.1 DO averages- days and nights- DAY
```{r 11.1 DO averages- days and nights- DAY}
## 11.1 DO averages- days and nights

# Creating Values
# Day

allchgDay_DO <- subset(H2Ochemdf, 
                       period == "day" &  treatment == "allchange",
                      select = c(dateTime, actualDOmg))

curDay_DO <- subset(H2Ochemdf, 
                   period == "day" & treatment == "current",
                   select = c(dateTime, actualDOmg))

hitempDay_DO <- subset(H2Ochemdf, 
                      period == "day" & treatment == "hightemperature",
                      select = c(dateTime, actualDOmg))

```


#### 11.2 DO averages- days and nights- NIGHT
```{r  11.2 DO averages- days and nights- NIGHT}

# Creating Values
# Night
allchgNightDO <- subset(H2Ochemdf, 
                        period == "night" & treatment == "allchange",
                        select = c(dateTime, actualDOmg ))

curNightDO <- subset(H2Ochemdf, 
                     period == "night" & treatment == "current",
                     select = c(dateTime, actualDOmg ))

hitempNightDO <- subset(H2Ochemdf, 
                        period == "night" & treatment == "hightemperature",
                        select = c(dateTime, actualDOmg ))


```


#### 11.3 Review of Values
```{r 11.3 Review of Values}
# 
# # Review of values
# allchgDayDO
# allchgNightDO
# curDayDO
# curNightDO
# hitempDayDO
# hitempNightDO

```


# 12.0) DO Day/Night Summary
#### 12.1 DO Day/Night Summary
```{r 12.1) DO Day/Night Summary}

H2Ochemdf.DO.daynight.summary <- H2Ochemdf %>% group_by(treatment, period) %>%
  summarize(sd = sd(actualDOmg, na.rm = TRUE), 
            mean = mean(actualDOmg, na.rm = TRUE), 
            median = median(actualDOmg, na.rm = TRUE),
            IQR = IQR(actualDOmg, na.rm = TRUE),
            n = n()) %>%
  mutate(se = sd/sqrt(n)) %>%
  mutate(ci = se*1.96)



write.csv(H2Ochemdf.DO.daynight.summary, "2020.07.26_H2Ochemdf_DO_daynight_summary.csv")

```


#### 12.2 H2Ochemdf_daynight_summary.csv

```{r 12.2 2020.07.23_H2Ochemdf_daynight_summary.csv posted as a tribble}
         tibble::tribble(
                  ~treatment,   ~period,         ~sd,       ~mean,     ~median,        ~IQR,    ~n,         ~se,         ~ci,
                 "allchange", "D2Nramp", 0.695200512, 7.545730356, 7.619970071, 0.899993471,  166L, 0.053958024, 0.105757728,
                 "allchange",     "day", 0.899952106, 5.012562386, 4.710812752, 0.821111536, 4802L, 0.012986984, 0.025454489,
                 "allchange",   "night", 0.362502769, 8.373891705, 8.450480339, 0.245667755, 2842L, 0.006799849, 0.013327705,
                   "current", "D2Nramp",  0.56039428, 8.333783768, 8.422599835,  0.89581169,  135L,  0.04823106, 0.094532878,
                   "current",     "day", 1.061486643, 6.077167812, 5.674745839, 0.901018275, 3849L, 0.017109628, 0.033534872,
                   "current",   "night", 0.436911474, 8.798980496, 8.915053838, 0.259654413, 2285L, 0.009140087, 0.017914571,
           "hightemperature", "D2Nramp", 0.502553157, 8.186262632, 8.281050428, 0.699788169,   86L, 0.054191701, 0.106215733,
           "hightemperature",     "day", 0.816148093, 5.658071726,    5.505228,  0.62372574, 2607L, 0.015984478, 0.031329577,
           "hightemperature",   "night", 0.197004247, 8.575377023, 8.585656769, 0.204390747, 1501L, 0.005084933, 0.009966469
           )

```



#### 12.3 Summary Table Plots

```{r}


ggplot(H2Ochemdf, aes(treatment, actualDOmg)) +
            geom_jitter(color = "grey") +
            geom_boxplot(notch = TRUE, outlier.shape = NA, colour = "green") +
            geom_point(data = H2Ochemdf.DO.daynight.summary, aes(x=treatment, y=mean), size=5, color = "purple") +
            geom_errorbar(data = H2Ochemdf.DO.daynight.summary,
                          aes(x=treatment, y=mean, ymin=mean-sd, ymax=mean+sd),
                          color = "blue") +
            geom_errorbar(data = H2Ochemdf.DO.daynight.summary,
                          aes(x=treatment, y=mean, ymin=mean-ci, ymax=mean+ci),
                          colour = "red") +
            facet_wrap(~period) +
            ggtitle("All Treatments, Actual Dissolved Oxygen (mg) Across Periods") +
            theme_bw() 


```


# 13.0 pH, Water Chemistry pH probes


```{r}

names(H2Ochemdf)


```

#### 13.1 pH (All Treatments & All MOATs) Colored by Period 

```{r 13.1 pH (All Treatments & All MOATs) Colored by Period }

pHplot1 <- ggplot(H2Ochemdf, aes(x=dateTime, y=pH)) +
  geom_point(aes(colour=period, point=))   +
  ggtitle("pH (All Treatments & All MOATs) Colored by Period")

pHplot1


```


#### 13.2 pH (All Treatments & All MOATs) Colored by Period, Facet Wrap
```{r 13.2 pH (All Treatments & All MOATs) Colored by Period, Facet Wrap}


pHplot2 <- ggplot(H2Ochemdf, aes(x=dateTime, y=pH)) +
  geom_point(aes(colour=period, point=))   +
  facet_wrap(~treatment) +
  ggtitle("System pH (All Treatments & All MOATs) Colored by Period")

pHplot2


```


#### 13.3 System pH (All Treatments & All MOATs) Day Period, subseted by period

```{r}
ggplot(subset(H2Ochemdf, 
               period %in% ("day")), 
        aes(x=dateTime, y=pH)) + 
   geom_point(aes(colour=treatment, point=)) +
   ylim (7.25, 7.80) + 
   ggtitle("System pH (All Treatments & All MOATs) Day Period, subseted by period")

```


#### 13.4 pH Day/Night Summary grouped by Period

```{r}
H2Ochemdf.pH.daynight.summary <- H2Ochemdf %>% group_by(treatment, period) %>%
  summarize(sd = sd(pH, na.rm = TRUE), 
            mean = mean(pH, na.rm = TRUE), 
            median = median(pH, na.rm = TRUE),
            IQR = IQR(pH, na.rm = TRUE),
            n = n()) %>%
  mutate(se = sd/sqrt(n)) %>%
  mutate(ci = se*1.96)



write.csv(H2Ochemdf.pH.daynight.summary, "2020.07.26_H2Ochemdf_pH_daynight_summary.csv")
```


#### 13.4a pH Day/Night Summary grouped by Period Table/Tribble
```{r}
tibble::tribble(
         ~treatment,   ~period,         ~sd,       ~mean,     ~median,        ~IQR,    ~n,         ~se,         ~ci,
        "allchange", "D2Nramp",  0.02390936, 7.573552387,    7.574293, 0.029397125,  166L, 0.001855726, 0.003637223,
        "allchange",     "day", 0.029965137, 7.532944951,  7.52721125, 0.031166938, 4802L, 0.000432419, 0.000847542,
        "allchange",   "night", 0.021372153, 7.561789777, 7.562601875, 0.031900687, 2842L,   0.0004009, 0.000785764,
          "current", "D2Nramp", 0.033368568, 7.708871117,   7.7186095,  0.02729375,  135L, 0.002871909, 0.005628942,
          "current",     "day", 0.035810953, 7.656779556,  7.65219375,  0.02552525, 3849L, 0.000577221, 0.001131353,
          "current",   "night", 0.044139727, 7.704555758,   7.7163225,   0.0269645, 2285L, 0.000923393,  0.00180985,
  "hightemperature", "D2Nramp", 0.020169313, 7.717802166, 7.720068625, 0.017174688,   86L, 0.002174913, 0.004262829,
  "hightemperature",     "day", 0.030866251, 7.661411591,  7.65717825,  0.02792625, 2607L, 0.000604524, 0.001184867,
  "hightemperature",   "night",  0.01825131,  7.72412941,     7.72807,  0.01511475, 1501L,  0.00047109, 0.000923336
  )

```





#### 13.5 pH Day/Night Summary Plot

```{r}
ggplot(H2Ochemdf, aes(treatment, pH)) +
            geom_jitter(color = "grey") +
            geom_boxplot(notch = TRUE, outlier.shape = NA, colour = "green") +
            geom_point(data = H2Ochemdf.pH.daynight.summary, aes(x=treatment, y=mean), size=5, color = "purple") +
            geom_errorbar(data = H2Ochemdf.pH.daynight.summary,
                          aes(x=treatment, y=mean, ymin=mean-sd, ymax=mean+sd),
                          color = "blue") +
            geom_errorbar(data = H2Ochemdf.pH.daynight.summary,
                          aes(x=treatment, y=mean, ymin=mean-ci, ymax=mean+ci),
                          colour = "red") +
            facet_wrap(~period) +
            ggtitle("All Treatments, pH Across Periods") +
            theme_bw() 
```
























